---
title: "Emotional salience modulates the forward flow of memory"
author: "Peris-Yague et al. 2021"
date: "8/17/2021"
output: 
  html_document:
    df_print: paged
  github_document:
    toc: true
  pdf_document: default
  word_document: default
---

```{r  echo=FALSE, results='hide', message=FALSE, warning=FALSE}
chooseCRANmirror(graphics=FALSE, ind=1)
knitr::opts_chunk$set(echo = TRUE)

#Install all libraries 

library(tidyverse)
library(ggpubr)
library(rstatix)
library(rsample)
library(dplyr)
library(ggplot2)
library(data.table)
library(effects)
library(lme4)
library(plyr)
library(emmeans)
library(reshape2)
```

# ***Manuscript Figures and Analyses***
# Fig. 1C- Amount of items recalled per SOA

```{r echo=FALSE,warning=FALSE, message=FALSE}
setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP_Code_Reviewed/PerisYague2023/Raw_Results")

listrecallSOA<-read.csv('list_recall_SOA_R.csv')
listrecallSOA<-as.data.frame(listrecallSOA)
listrecallSOA$SOA<-factor(listrecallSOA$SOA)
listrecallSOA$subject<-factor(listrecallSOA$subject)

listrecallSOA.data<- listrecallSOA %>%
  group_by(SOA,subject) %>%
  get_summary_stats(total_recall_list, type='mean')

listrecallSOA.data<-listrecallSOA.data[-c(3,4)]
names(listrecallSOA.data)[3] <- "total_recall_list"

listrecallSOA.data<-as.data.frame(listrecallSOA.data)
listrecallSOA.data$SOA<-as.factor(listrecallSOA.data$SOA)
listrecallSOA.data$subject<-as.factor(listrecallSOA.data$subject)
listrecallSOA.data$total_recall_list<-as.numeric(listrecallSOA.data$total_recall_list)
  
listrecallSOA.wide<-reshape(listrecallSOA, idvar = 'subject', timevar = 'SOA', direction = 'wide')
names(listrecallSOA.wide)[2] <- "SOA 6"
names(listrecallSOA.wide)[3] <- "SOA 3"
names(listrecallSOA.wide)[4] <- "SOA 2"
names(listrecallSOA.wide)[5] <- "SOA 1"
names(listrecallSOA.wide)[6] <- "SOA 4"


## Plot recall performance by SOA in % 
per100recallbySOA <- listrecallSOA
per100recallbySOA$total_recall_list<-(per100recallbySOA$total_recall_list)/14*100 #Now we have the % of items recalled

listrecallSOAplot2<-per100recallbySOA %>%
  group_by(SOA) %>%
  get_summary_stats(total_recall_list,type='mean_ci')

per100recallbySOA <-per100recallbySOA %>%
  group_by(SOA, subject) %>%
  get_summary_stats(total_recall_list, type='mean_ci')

# Plot 
list.per100.recall.SOA.fig2<-ggplot(data=listrecallSOAplot2, aes(x=as.factor(SOA), y=mean, color=SOA, ymin=mean-ci, ymax=mean+ci))+
  #geom_pointrange(aes(ymin=mean-ci,ymax=mean+ci) ,position=position_dodge(width=0.5), shape=15, size=0.5)+
  geom_errorbar(aes(x=SOA, ymin=mean-ci, ymax=mean+ci), width=0.4, size=1, position=position_dodge(0.9), stat='identity')+
    geom_point(position=position_dodge(0.9), size=3, shape=20) +
  geom_jitter(data=per100recallbySOA,aes(color=SOA, x=SOA, y=mean), position=position_jitterdodge(dodge.width = 0,2),  show.legend=FALSE, alpha=0.3, inherit.aes=F)+
  scale_color_brewer(palette="Set2")+
  scale_y_continuous(n.breaks=10)+
   #geom_hline(yintercept=50, linetype = 'dashed', color = 'grey')+
  theme_classic()+
  labs(x='Stimulus Onset Asynchrony', y='% items recalled')+
  theme(text=element_text(size=25))+
  scale_y_continuous(
  labels = scales::number_format(accuracy = 1))

list.per100.recall.SOA.fig2

```


# Fig. 1D- Oddballs vs controls output position
```{r echo=FALSE, warning=FALSE, message=FALSE}
setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP_Code_Reviewed/PerisYague2023/Raw_Results")

oddballs<-read.csv('oddlists.csv')
oddballs$enc_pos <- as.numeric(oddballs$enc_pos)
oddballs$rec_pos <- as.numeric(oddballs$rec_pos)
oddballs$list_type <- as.factor(oddballs$list_type)
oddballs$subject <- as.factor(oddballs$subject)
oddballs <- drop_na(oddballs) #drop the trials where odd not recalled

controls <- read.csv('ctrlists.csv')
controls$enc_pos <- as.numeric(controls$enc_pos)
controls$rec_pos <- as.numeric(controls$rec_pos)
controls$list_type <- as.factor(controls$list_type)
controls$subject <- as.factor(controls$subject)
controls <- drop_na(controls) #drop the trials where controls not recalled

#Only take trials where the oddballs and controls were encoded in the 7,8,11,12th positions
odd_filt <- filter(oddballs, enc_pos == '7' | enc_pos == '8' | enc_pos == '11' | enc_pos == '12')
ctrl_filt <- filter(controls, enc_pos == '7' | enc_pos == '8' | enc_pos == '11' | enc_pos == '12')

# Plot them sepparately 
# Check out oddballs
plt<-ggplot(odd_filt, aes(fill=list_type, x=enc_pos, y=rec_pos, color=list_type))+
  geom_jitter()+
  scale_y_continuous(breaks=c(0,1,2,3,4,5,6,7, 8, 9, 10, 11, 12, 13, 14))+
  scale_x_continuous(breaks=c(7,8,11,12))

# Check out controls 
plt<-ggplot(ctrl_filt, aes(fill=list_type, x=enc_pos, y=rec_pos, color=list_type))+
  geom_jitter()+
  scale_y_continuous(breaks=c(0,1,2,3,4,5,6,7, 8, 9, 10, 11, 12, 13, 14))+
  scale_x_continuous(breaks=c(7,8,11,12))

data <- rbind(ctrl_filt, odd_filt)
data_summary <- data %>%
  group_by(list_type,item) %>%
  get_summary_stats(rec_pos, type='mean_ci')

# data for jitter adding one dot per subject 
data.jitter <- data %>% 
  group_by(list_type,item, subject) %>%
  get_summary_stats(rec_pos, type = 'mean_ci')

plot <- ggplot(data_summary, aes(fill=list_type, x=item, y=mean))+
  geom_bar(position='dodge', stat='identity')+
  geom_errorbar(aes(x=item, ymin=mean-ci, ymax=mean+ci), width=0.4, color='wheat4', size=0.5, position=position_dodge(0.9), stat='identity')+
  geom_jitter(data=data.jitter,aes(color=list_type, x=item, y=mean), position=position_jitterdodge(dodge.width = 0.9), show.legend=FALSE, alpha=0.2)+
  guides(col=FALSE)+
  scale_color_manual(values=c('orchid', 'deepskyblue'))+
  scale_fill_manual(values=c('thistle1', 'lightblue1'), labels = c('Emotional', 'Perceptual'), name= 'List Type')+
  scale_x_discrete(labels=c('Control', 'Oddball'))+
  scale_y_continuous(breaks=c(1,2,3,4,5,6,7, 8, 9, 10, 11, 12, 13, 14))+
  labs(x='Item type', y='Output position')+
  theme_classic()+
  theme(text=element_text(size=25))
  

plot

data_stats<- data %>%
  group_by(subject,item,list_type) %>%
  get_summary_stats(rec_pos, type='mean_ci')

data_stats$mean <- as.numeric(data_stats$mean)
data_stats$subject <- as.factor(data_stats$subject)
data_stats$item <- as.factor(data_stats$item)
data_stats$list_type <-as.factor(data_stats$list_type)

output.QQ<-ggqqplot(data_stats,'mean',ggtheme=theme_bw())+
  facet_grid(list_type~item, labeller='label_both')

data_stats %>% levene_test(mean ~ list_type*item)

data_aov<-anova_test(data=data_stats, dv=mean, wid=subject,
                 within=c(list_type,item), effect.size = 'pes',
                 )
get_anova_table(data_aov) 


#write.csv(data_stats,"/output.csv", row.names = FALSE)


```


# Fig. 2B Conditional Response Probability Curves


```{r echo=FALSE,warning=FALSE, message=FALSE}
setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP_Code_Reviewed/PerisYague2023/Raw_Results")
CRPall<-read.csv('CRP_all_R.csv')
CRPall<-as.data.frame(CRPall)

crpplot<-CRPall %>%
  group_by(oddballtype,wordposition, direction) %>%
  get_summary_stats(CRP,type='mean_ci')

## This is needed to make the 'backwards' lags negative so that they plot
# correctly 
wpplot<-numeric(0)
for (rows in 1:nrow(crpplot)){
  if (crpplot$direction[rows] == "Backwards") {
    wpplot[rows]<-(crpplot$wordposition[rows])*-1
  }
  else {wpplot[rows]<-(crpplot$wordposition[rows])
  }
}
invisible(as.data.frame(wpplot))
crpplot$wpplot<-wpplot

pltCRP<-ggplot(crpplot,aes(x=wpplot,y=mean), group=oddballtype)+
        geom_line(aes(color=oddballtype), size=1) +
        geom_point(aes(color=oddballtype), size=2,) +
        geom_ribbon(aes(x=wpplot, ymin=mean-ci, ymax=mean+ci, fill=oddballtype), alpha=0.3, show.legend = FALSE)+
        #geom_errorbar(aes(ymin=mean-ci, ymax=mean+ci,color=oddballtype), alpha=0.5)+
        theme_classic()+
        facet_wrap(~ direction, scales='free_x', ncol=2)+    
        theme(strip.background = element_blank())+
        theme(text=element_text(size=15))+#25))+
        #ggtitle('CRP all')+
        xlab('Lag') +
        ylab('Conditional Response Probability')+
        scale_color_manual(values=c('violetred2','turquoise2'),aesthetics = c('colour','fill'), 
        labels=c('Emotional','Perceptual'), name='Oddball List Type')+
        scale_y_continuous(labels = scales::number_format(accuracy = 0.01))
                    
pltCRP

```




# Fig. 2D- 2E To vs from transitions from oddball items

```{r echo=FALSE, warning=FALSE, message=FALSE}

# Transitions from oddballs 
setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP_Code_Reviewed/PerisYague2023/Raw_Results")
TFCRP<-read.csv('ToFrom_CRP_R.csv')
TFCRP<-as.data.frame(TFCRP)
#head(TFCRP) #to show you what my raw data looks like 

TFplot<-TFCRP %>%
  group_by(oddballtype,wordposition, direction,transition) %>%
  get_summary_stats(CRP,type='mean_ci')

wpplot<-numeric(0)
for (rows in 1:nrow(TFplot)){
  if (TFplot$direction[rows] == "Backwards") {
    wpplot[rows]<-(TFplot$wordposition[rows])*-1
  }
  else {wpplot[rows]<-(TFplot$wordposition[rows])
  }
}
invisible(as.data.frame(wpplot))
TFplot$wpplot<-wpplot
TFplot$transition<-factor(TFplot$transition, levels=c('to','from'))

transition.labs<- c('Transition TO the oddball', 'Transition FROM the oddball')
names(transition.labs)<-c('to', 'from')

TFplot <- filter(TFplot, transition == 'from') 
TFplot <- droplevels(TFplot)

# Transitions from controls 
setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP/Raw_Results")
TFCRPctrl<-read.csv('control_ToFrom_CRP_R.csv')
TFCRPctrl<-as.data.frame(TFCRPctrl)
#head(TFCRP) #to show you what my raw data looks like 

ctrlTFplot<-TFCRPctrl %>%
  group_by(oddballtype,wordposition, direction,transition) %>%
  get_summary_stats(CRP,type='mean_ci')

ctrlwpplot<-numeric(0)
for (rows in 1:nrow(ctrlTFplot)){
  if (ctrlTFplot$direction[rows] == "Backwards") {
    ctrlwpplot[rows]<-(ctrlTFplot$wordposition[rows])*-1
  }
  else {ctrlwpplot[rows]<-(ctrlTFplot$wordposition[rows])
  }
}
invisible(as.data.frame(ctrlwpplot))

ctrlTFplot$ctrlwpplot<-ctrlwpplot
ctrlTFplot$transition<-factor(ctrlTFplot$transition, levels=c('to','from'))

ctrlTFplot <- filter(ctrlTFplot, transition == 'from')
ctrlTFplot <- droplevels(ctrlTFplot)

# Rename the from to oddball / control to concatenate the lists for plotting 
TFplot$transition <- as.character(TFplot$transition)
TFplot["transition"][TFplot["transition"] == "from"] <- "oddball"
TFplot$transition <- as.factor(TFplot$transition)

ctrlTFplot$transition <- as.character(ctrlTFplot$transition)
ctrlTFplot["transition"][ctrlTFplot["transition"] == "from"] <- "control"
names(ctrlTFplot)[names(ctrlTFplot) == "ctrlwpplot"] <- "wpplot"

ctrlTFplot$transition <- as.factor(ctrlTFplot$transition)

data <- rbind(TFplot, ctrlTFplot)

transition.labs<- c('Transition from oddballs', 'Transition from controls')
names(transition.labs)<-c('oddball', 'control')


pltTF<-ggplot(data,aes(x=wpplot,y=mean), group=oddballtype)+
  geom_line(aes(color=oddballtype), size=1) +
  geom_point(aes(color=oddballtype), size=2) +
  geom_ribbon(aes(ymin=mean-ci, ymax=mean+ci, fill=oddballtype), alpha=0.2, show.legend = FALSE, inherit.aes=TRUE)+
  #geom_errorbar(aes(ymin=mean-se, ymax=mean+se,color=oddballtype), alpha=0.5)+
   theme_classic()+
  facet_grid(transition~direction, scales='free_x', 
            labeller=labeller(transition = as_labeller(transition.labs)))+   
  theme(strip.background = element_blank())+
  theme(text=element_text(size=15))+#25))+
  xlab('Lag') +
  ylab('Conditional Response Probability')+
  scale_color_manual(values=c('violetred2','turquoise2'), 
                     labels=c('Emotional','Perceptual'), name='Oddball List Type', aesthetics=c('colour','fill'))+
  scale_y_continuous(
  labels = scales::number_format(accuracy = 0.01))
 

pltTF


#Compare the averaged (+- 5 lags) transitions to and from the oddballs vs. the control items 
CRP_odd_v_ctrl<-read.csv('CRP_avgd_oddballvcontrol.csv')
CRP_odd_v_ctrl<-as.data.frame(CRP_odd_v_ctrl)

CRP_odd_v_ctrl$subject<-as.factor(CRP_odd_v_ctrl$subject)
CRP_odd_v_ctrl$oddballtype<-as.factor(CRP_odd_v_ctrl$oddballtype)
CRP_odd_v_ctrl$transition<-as.factor(CRP_odd_v_ctrl$transition)
CRP_odd_v_ctrl$direction<-as.factor(CRP_odd_v_ctrl$direction)
CRP_odd_v_ctrl$item_type<-as.factor(CRP_odd_v_ctrl$item_type)

transition.labs<- c('Transitions from the target items')
names(transition.labs)<-c('from')

CRP_odd_v_ctrl <- filter(CRP_odd_v_ctrl, transition == 'from')
CRP_odd_v_ctrl <- droplevels(CRP_odd_v_ctrl)

########## 3 way ANOVA #
CRP_odd_v_ctrl$CRPmean <- as.numeric(CRP_odd_v_ctrl$CRPmean)
CRP_odd_v_ctrl$direction <- as.factor(CRP_odd_v_ctrl$direction)
CRP_odd_v_ctrl$oddballtype <- as.factor(CRP_odd_v_ctrl$oddballtype)
CRP_odd_v_ctrl$item_type <- as.factor(CRP_odd_v_ctrl$item_type)
  
#ggqqplot(CRP_odd_v_ctrl,'CRPmean',ggtheme=theme_bw())+
 # facet_grid(oddballtype~item_type+direction, labeller='label_both')

#crp.3way <- anova_test(data=CRP_odd_v_ctrl, dv=CRPmean, wid=subject,
  #               within=c(oddballtype,item_type,direction), effect.size = 'pes'
 #                )
#get_anova_table(crp.3way) 

#post-hoc tests
#CRP_odd_v_ctrl %>%
 # group_by(oddballtype,direction) %>%
#  pairwise_t_test(CRPmean~item_type, paired=TRUE, p.adjust.method='fdr')

#CRP_odd_v_ctrl %>%
  #group_by(oddballtype,direction) %>%
 # cohens_d(CRPmean~item_type, paired=TRUE)

###

plt <- CRP_odd_v_ctrl %>% 
  group_by(oddballtype, item_type, transition, direction) %>%
  get_summary_stats(CRPmean, type = 'mean_ci')

pltTF.collapsed<-ggplot(plt,aes(x=direction,y=mean, group=oddballtype, shape= item_type, color = oddballtype,  ymin=mean-ci, ymax=mean+ci))+
    geom_pointrange(position=position_dodge2(width=0.5), size=1)+
   theme_classic()+
  facet_grid(~transition, scales='free_x', 
             labeller=labeller(transition = as_labeller(transition.labs)))+   
  theme(text=element_text(size=15))+#25))+
  theme(strip.background = element_blank())+
  xlab('Collapsed Lag (± 5 lags)') +
  ylab('Conditional Response Probability')+
  scale_color_manual(values=c('violetred2','turquoise2'), 
                     labels=c('Emotional','Perceptual'), name='Oddball List Type')+
  scale_shape_manual(name = "Item Type",
                     labels = c("Control", "Oddball"),
                     values=c(16,17)) 

  pltTF.collapsed
  
  # Run statistics
  fwrd <- CRP_odd_v_ctrl
  fwrd$CRPmean <- as.numeric(fwrd$CRPmean)
  # only select forward transitions as this is what we're interested in 
  fwrd <- filter(fwrd, direction == 'forwards')
  #
  fwrd$direction <- as.factor(fwrd$direction)
  fwrd$oddballtype <- as.factor(fwrd$oddballtype)
  fwrd$item_type <- as.factor(fwrd$item_type)
  
  fwrd <- droplevels(fwrd)
  
 #ggqqplot(fwrd,'CRPmean',ggtheme=theme_bw())+
  #facet_grid(oddballtype~item_type, labeller='label_both')

transitions_aov<-anova_test(data=fwrd, dv=CRPmean, wid=subject,
                 within=c(oddballtype,item_type), effect.size = 'pes'
                 )
get_anova_table(transitions_aov) 


fwrd %>%
  group_by(oddballtype) %>%
  pairwise_t_test(CRPmean~item_type, paired=TRUE, p.adjust.method='fdr')

fwrd %>%
  group_by(oddballtype) %>%
  cohens_d(CRPmean~item_type, paired=TRUE)

# For completion check backward transitions
bkwrd <- CRP_odd_v_ctrl 
bkwrd$CRPmean <- as.numeric(bkwrd$CRPmean)
  # only select forward transitions as this is what we're interested in 
bkwrd <- filter(bkwrd, direction == 'backwards')
  #
bkwrd$direction <- as.factor(bkwrd$direction)
bkwrd$oddballtype <- as.factor(bkwrd$oddballtype)
bkwrd$item_type <- as.factor(bkwrd$item_type)
  
bkwrd <- droplevels(bkwrd)
  
#ggqqplot(bkwrd,'CRPmean',ggtheme=theme_bw())+
 # facet_grid(oddballtype~item_type, labeller='label_both')

bckwrd_transitions_aov<-anova_test(data=bkwrd, dv=CRPmean, wid=subject,
                 within=c(oddballtype,item_type), effect.size = 'pes'
                 )
get_anova_table(bckwrd_transitions_aov) 


bkwrd %>%
  group_by(oddballtype) %>%
  pairwise_t_test(CRPmean~item_type, paired=TRUE, p.adjust.method='fdr')

bkwrd %>%
  group_by(oddballtype) %>%
  cohens_d(CRPmean~item_type, paired=TRUE)



```


# ***Supplementary Material Figures and Analyses***
# Fig. S1A,S1B, Table S1- Recall probability
```{r echo=FALSE, warning=FALSE, message=FALSE}
setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP_Code_Reviewed/PerisYague2023/Raw_Results")

data<-read.csv('recall_probability_R.csv')
data$recprob <- as.numeric(data$recprob)
data$subject <- as.factor(data$subject)
data$encpos <- as.factor(data$encpos)
data$list_type <- as.factor(data$list_type)
data$item <- as.factor(data$item)
data$oddvctrl <- as.factor(data$oddvctrl)
data <- drop_na(data)
data <- droplevels(data)

# Figure S1a

#summarize data for plot
data.plt <- data %>%
  group_by(item,list_type,encpos) %>%
  get_summary_stats(recprob,type='mean_ci')

#Let's visualize it 
#Plot oddballs vs. controls
data.plt3 <- data %>%
  group_by(item,list_type) %>%
  get_summary_stats(recprob,type='mean_ci')

data.plt3 <- data.plt3[ which(data.plt3$item=='control'| data.plt3$item=='oddball'), ]
data.plt3 <-droplevels(data.plt3)

data.plt3.jitter <- data %>%
  group_by(item,list_type,subject) %>%
  get_summary_stats(recprob, type= 'mean_ci')

data.plt3.jitter <- data.plt3.jitter[which(data.plt3.jitter$item == 'control' | data.plt3.jitter$item == 'oddball'),]

plot3<-ggplot(data.plt3, aes(fill=list_type, x=item, y=mean))+
  geom_bar(position='dodge', stat='identity')+
  geom_errorbar(aes(x=item, ymin=mean-ci, ymax=mean+ci), width=0.4, color='wheat4',size=0.5,
                position=position_dodge(0.9), stat='identity')+
  geom_jitter(data=data.plt3.jitter, aes(color=list_type,x=item, y=mean), position=position_jitterdodge(dodge.width = 0.9),show.legend = FALSE, alpha=0.2)+
  guides(col=FALSE)+
  scale_color_manual(values=c('orchid','deepskyblue'))+
  scale_fill_manual(values=c('thistle1','lightblue1'),labels=c('Emotional', 'Perceptual'), name= 'List type')+
 scale_x_discrete(labels=c('Control','Oddball'))+
  labs(x='Item type', y='Recall probability')+
  theme_classic()+
  theme(strip.background=element_blank(), text=element_text(size=17))

plot3

# Run an ANOVA 
data.plt3.jitter <-as.data.frame(data.plt3.jitter)
data.plt3.jitter$subject <- as.factor(data.plt3.jitter$subject)
data.plt3.jitter$list_type <- as.factor(data.plt3.jitter$list_type)
data.plt3.jitter$item <- as.factor(data.plt3.jitter$item)
data.plt3.jitter$mean <- as.numeric(data.plt3.jitter$mean)

# check for normality 
plt3.QQ<-ggqqplot(data.plt3.jitter,'mean',ggtheme=theme_bw())+
  facet_grid(list_type~item, labeller='label_both')

data.plt3.jitter %>% levene_test(mean ~ list_type*item)

plt3_aov<-anova_test(data=data.plt3.jitter, dv=mean, wid=subject,
                 within=c(list_type,item),
                 effect.size='pes')
get_anova_table(plt3_aov) 

# post-hoc test for item type 
data.plt3.jitter <- droplevels(data.plt3.jitter)

data.plt3.jitter %>%
  group_by(item,subject) %>%
  get_summary_stats(mean, type= 'mean_ci') %>%
  pairwise_t_test(mean~item, paired=TRUE, p.adjust.method='fdr')

data.plt3.jitter %>%
  group_by(item,subject) %>%
  get_summary_stats(mean, type= 'mean_ci') %>%
  cohens_d(mean~item, paired=TRUE)

#post-hoc for list type
data.plt3.jitter %>%
  group_by(list_type,subject) %>%
  get_summary_stats(mean, type= 'mean_ci') %>%
  pairwise_t_test(mean~list_type, paired=TRUE, p.adjust.method='fdr')

data.plt3.jitter %>%
  group_by(list_type,subject) %>%
  get_summary_stats(mean, type= 'mean_ci') %>%
  cohens_d(mean~list_type, paired=TRUE)


# Table S1, Fig. S1B, Plot items to test retrograde amnesic effect 
data.plt4 <- data %>%
  group_by(item,list_type) %>%
  get_summary_stats(recprob,type='mean_ci')

data.plt4 <- data.plt4[ which(data.plt4$item =='controlm2'| data.plt4$item =='controlm1' | data.plt4$item == 'controlp1'
                              |data.plt4$item == 'oddballm2' | data.plt4$item == 'oddballm1' | data.plt4$item == 'oddballp1'), ]

data.plt4 <- droplevels(data.plt4)

data.plt4.jitter <- data %>%
  group_by(item,list_type,subject,oddvctrl) %>%
  get_summary_stats(recprob, type= 'mean_ci')

data.plt4.jitter <- data.plt4.jitter[ which(data.plt4.jitter$item =='controlm2'| data.plt4.jitter$item =='controlm1' |
                                              data.plt4.jitter$item == 'controlp1'|data.plt4.jitter$item == 'oddballm2' | 
                                              data.plt4.jitter$item == 'oddballm1' | data.plt4.jitter$item == 'oddballp1'), ]


plot4<-ggplot(data.plt4, aes(fill=list_type, x=item, y=mean))+
  geom_bar(position='dodge', stat='identity')+
  geom_errorbar(aes(x=item, ymin=mean-ci, ymax=mean+ci), width=0.4, color='wheat4',size=0.5,
                position=position_dodge(0.9), stat='identity')+
  geom_jitter(data=data.plt4.jitter, aes(color=list_type,x=item, y=mean), position=position_jitterdodge(dodge.width = 0.9),show.legend = FALSE, alpha=0.2)+
  guides(col=FALSE)+
  scale_color_manual(values=c('orchid','deepskyblue'))+
  scale_fill_manual(values=c('thistle1','lightblue1'),labels=c('Emotional', 'Perceptual'), name= 'List type')+
 scale_x_discrete(labels=c('Ctrl-1','Ctrl-2', 'Ctrl+1', 'Odd-1', 'Odd-2', 'Odd+1'))+
  labs(x='Item type', y='Recall probability')+
  theme_classic()+
  theme(strip.background=element_blank(), text=element_text(size=17))

plot4
  

data.plt4.jitter <- droplevels(data.plt4.jitter)
data.plt4.jitter$item <- as.character(data.plt4.jitter$item)

data.plt4.jitter["item"][data.plt4.jitter["item"] == "controlm1"] <- "m1"
data.plt4.jitter["item"][data.plt4.jitter["item"] == "controlm2"] <- "m2"
data.plt4.jitter["item"][data.plt4.jitter["item"] == "controlp1"] <- "p1"
data.plt4.jitter["item"][data.plt4.jitter["item"] == "oddballm2"] <- "m2"
data.plt4.jitter["item"][data.plt4.jitter["item"] == "oddballm1"] <- "m1"
data.plt4.jitter["item"][data.plt4.jitter["item"] == "oddballp1"] <- "p1"

data.plt4.jitter$item <- as.factor(data.plt4.jitter$item)
data.plt4.jitter$oddvctrl <- as.factor(data.plt4.jitter$oddvctrl)


plt4.QQ<-ggqqplot(data.plt4.jitter,'mean',ggtheme=theme_bw())+
  facet_grid(list_type~item*oddvctrl, labeller='label_both')

data.plt4.jitter %>% levene_test(mean ~ list_type*item*oddvctrl)

plt4_aov <- anova_test(mean ~ list_type*item*oddvctrl + Error(subject/(list_type*item*oddvctrl)), data=data.plt4.jitter)
  
plt4_aov<-anova_test(
  data=data.plt4.jitter,dv=mean,wid=subject, 
  within=c(list_type, oddvctrl,item), type=3, effect.size = 'pes'
)

get_anova_table(plt4_aov)

data.plt4.jitter %>%
  group_by(list_type,subject) %>%
  get_summary_stats(mean, type= 'mean_ci') %>%
  pairwise_t_test(mean~list_type, paired=TRUE, p.adjust.method='fdr')

data.plt4.jitter %>%
  group_by(list_type,subject) %>%
  get_summary_stats(mean, type= 'mean_ci') %>%
  cohens_d(mean~list_type, paired=TRUE)

# Group the data by encoding position 
datajasp <-data %>%
  group_by(subject,list_type,item) %>%
  get_summary_stats(recprob,type='mean_ci')
  
#go to jasp
test.wide <- dcast(datajasp, subject ~  list_type + item, value.var = "mean", drop = FALSE)

path_out<- '/Raw_Results/'

#write.csv(test.wide, paste(path_out,'recallprobability_JASP.csv',sep = ''))
#####

```


# Fig. S2 Conditional Response Probability Curves by Stimulus Onset Asynchrony

```{r echo=FALSE, warning=FALSE, message=FALSE}
setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP_Code_Reviewed/PerisYague2023/Raw_Results")
CRPSOA<-read.csv('all_CRP_SOA_R.csv')
CRPSOA<-as.data.frame(CRPSOA)

CRPSOAplot<-CRPSOA %>%
  group_by(wordposition, direction, SOA) %>%
  get_summary_stats(CRP,type='mean_ci')

wpplot<-numeric(0)
for (rows in 1:nrow(CRPSOAplot)){
  if (CRPSOAplot$direction[rows] == "Backwards") {
    wpplot[rows]<-(CRPSOAplot$wordposition[rows])*-1
  }
  else {wpplot[rows]<-(CRPSOAplot$wordposition[rows])
  }
}
invisible(as.data.frame(wpplot))
CRPSOAplot$wpplot<-wpplot

#factor all the categorical variables 
CRPSOA$SOA<-factor(CRPSOA$SOA)
CRPSOA$direction<-factor(CRPSOA$direction)
CRPSOA$wordposition<-factor(CRPSOA$wordposition)
CRPSOA$subject<-factor(CRPSOA$subject)
CRPSOA$oddballtype<-factor(CRPSOA$oddballtype)

pltCRPSOA<-ggplot(CRPSOAplot,aes(x=wpplot,y=mean, group=SOA, color=as.factor(SOA)))+
  geom_line(size=1) +
  geom_point(size=2) +
  geom_ribbon(aes(ymin=mean-ci, ymax=mean+ci, fill=as.factor(SOA)), alpha=0.2, show.legend = FALSE, inherit.aes=TRUE)+
  #geom_errorbar(aes(ymin=mean-ci, ymax=mean+ci,color=as.factor(SOA)), alpha=0.5)+
  theme_classic()+
  facet_wrap(~ direction, scales='free_x', ncol=2)+ 
  theme(strip.background = element_blank())+
  theme(text=element_text(size=17))+
  xlab('Lag') +
  ylab('Conditional Response Probability')+
  scale_color_brewer(palette="Set2", aesthetics = c('colour','fill') )+
  labs(color='SOA')+
  scale_y_continuous(
  labels = scales::number_format(accuracy = 0.01))

pltCRPSOA

  
```


# Fig. S3- Correlation of lag + 1 from emotional oddballs with E-1 recall 
```{r echo=FALSE, warning=FALSE, message=FALSE}

# Can easily do a correlation on a subject by subject basis  ?? ¿¿ 
em1_corr <- as.data.frame(subset(data.plt4.jitter, list_type == 'emotional' & item =='m1' & oddvctrl=='oddball')) #raw recall of E-1 items

lag_corr <- as.data.frame(subset(CRP_odd_v_ctrl, oddballtype=='E' & transition == 'from' & direction=='forwards' & item_type == 'oddball'))

corrplot<-as.data.frame(cbind(em1_corr$mean,lag_corr$CRPmean))

corrplt<-ggplot(corrplot, aes(x=V1, y=V2))+
  geom_point(shape=18, color='darksalmon')+
  geom_smooth(method=lm, se=FALSE, color='lightskyblue2')+
  #ggtitle('Correlation of E-1 normalized recall and transitions from E Odd (lag +1)')+
  ylab('Lag +1 transitions from E Odd')+
  xlab('E-1 Recall')+
  theme_classic()+
  theme(strip.background = element_blank())+
  theme(text=element_text(size=17))

corrplt

cor.test(corrplot$V1, corrplot$V2,  method = "spearman", use = "complete.obs")
``` 

# Output for JASP in table S3, S4, S5- recall probability by SOA (view code)
```{r echo=FALSE, warning=FALSE, message=FALSE}
setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP_Code_Reviewed/PerisYague2023/Raw_Results")

data<-read.csv('recall_probability_SOA_R.csv')
data$recprob <- as.numeric(data$recprob)
data$subject <- as.factor(data$subject)
data$encpos <- as.factor(data$encpos)
data$list_type <- as.factor(data$list_type)
data$item <- as.factor(data$item)
data$soa <- as.factor(data$soa)

data <- drop_na(data)

data.plt <- data %>%
  group_by(item,list_type,soa) %>%
  get_summary_stats(recprob,type='mean_ci')

#Let's visualize it 
plot <- ggplot(data.plt, aes(fill=list_type, x=item, y=mean))+
  facet_wrap(~soa)+
  geom_bar(position='dodge', stat='identity')+
  geom_errorbar(aes(x=item, ymin=mean-ci, ymax=mean+ci), width=0.4, color='wheat4',size=0.5,
                position=position_dodge(0.9), stat='identity')+
  geom_jitter(data=data, aes(color=list_type,x=item, y=recprob), position=position_jitterdodge(dodge.width = 0.9),show.legend = FALSE, alpha=0.2)+
  guides(col=FALSE)+
  scale_color_manual(values=c('orchid','deepskyblue'))+
  scale_fill_manual(values=c('thistle1','lightblue1'),labels=c('Emotional', 'Perceptual'), name= 'List type')+
 scale_x_discrete(labels=c('Control','Ctrl-1','Ctrl-2','Ctrl+1','Oddball','Odd-1','Odd-2','Odd+1'))+
  labs(x='Item type', y='Recall probability')+
  theme_classic()+
  theme(strip.background=element_blank(), text=element_text(size=17))

# Now let's plot oddballs vs. controls by SOA
data.plt3 <- data %>%
  group_by(item,list_type,soa) %>%
  get_summary_stats(recprob,type='mean_ci')

data.plt3 <- data.plt3[ which(data.plt3$item=='control'| data.plt3$item=='oddball'), ]

data.plt3.jitter <- data %>%
  group_by(item,list_type,soa,subject) %>%
  get_summary_stats(recprob, type= 'mean_ci')

data.plt3.jitter <- data.plt3.jitter[which(data.plt3.jitter$item == 'control' | data.plt3.jitter$item == 'oddball'),]

plot3<-ggplot(data.plt3, aes(fill=list_type, x=item, y=mean))+
  facet_wrap(~soa)+
  geom_bar(position='dodge', stat='identity')+
  geom_errorbar(aes(x=item, ymin=mean-ci, ymax=mean+ci), width=0.4, color='wheat4',size=0.5,
                position=position_dodge(0.9), stat='identity')+
  geom_jitter(data=data.plt3.jitter, aes(color=list_type,x=item, y=mean), position=position_jitterdodge(dodge.width = 0.9),show.legend = FALSE, alpha=0.2)+
  guides(col=FALSE)+
  scale_color_manual(values=c('orchid','deepskyblue'))+
  scale_fill_manual(values=c('thistle1','lightblue1'),labels=c('Emotional', 'Perceptual'), name= 'List type')+
 scale_x_discrete(labels=c('Control','Oddball'))+
  labs(x='Item type', y='Recall probability')+
  theme_classic()+
  theme(strip.background=element_blank(), text=element_text(size=17))

#plot3



# Retrograde amnesia effects by SOA 
data.plt4 <- data %>%
  group_by(item,list_type,soa) %>%
  get_summary_stats(recprob,type='mean_ci')

data.plt4 <- data.plt4[ which(data.plt4$item =='control-1'| data.plt4$item =='control-2' | data.plt4$item == 'control+1'
                              |data.plt4$item == 'oddball-1' | data.plt4$item == 'oddball-2' | data.plt4$item == 'oddball+1'), ]

data.plt4.jitter <- data %>%
  group_by(item,list_type,subject,soa) %>%
  get_summary_stats(recprob, type= 'mean_ci')

data.plt4.jitter <- data.plt4.jitter[ which(data.plt4.jitter$item =='control-1'| data.plt4.jitter$item =='control-2' |
                                              data.plt4.jitter$item == 'control+1'|data.plt4.jitter$item == 'oddball-1' | 
                                              data.plt4.jitter$item == 'oddball-2' | data.plt4.jitter$item == 'oddball+1'), ]


plot4<-ggplot(data.plt4, aes(fill=list_type, x=item, y=mean))+
  facet_wrap(~soa)+
  geom_bar(position='dodge', stat='identity')+
  geom_errorbar(aes(x=item, ymin=mean-ci, ymax=mean+ci), width=0.4, color='wheat4',size=0.5,
                position=position_dodge(0.9), stat='identity')+
  geom_jitter(data=data.plt4.jitter, aes(color=list_type,x=item, y=mean), position=position_jitterdodge(dodge.width = 0.9),show.legend = FALSE, alpha=0.2)+
  guides(col=FALSE)+
  scale_color_manual(values=c('orchid','deepskyblue'))+
  scale_fill_manual(values=c('thistle1','lightblue1'),labels=c('Emotional', 'Perceptual'), name= 'List type')+
 scale_x_discrete(labels=c('Control-1','Control-2', 'Control+1', 'Oddball-1', 'Oddball-2', 'Oddball+1'))+
  labs(x='Item type', y='Recall probability')+
  theme_classic()+
  theme(strip.background=element_blank(), text=element_text(size=17))

#plot4

#convert data to wide for JASP 
dat <-data %>%
  group_by(subject,list_type,item,soa) %>%
  get_summary_stats(recprob,type='mean_ci')

data.wide <- dcast(dat, subject ~  list_type + item + soa, value.var = "mean", drop = FALSE)

path_out<- '/Raw_Results/'

#write.csv(data.wide, paste(path_out,'recallprobability_SOA_JASP.csv',sep = ''))

#These results are inputted into a JASP file called 'recallprobability_SOA_JASP' for results of tables S3, S4, S5

```
# Table S6- Emotional ratings  

``` {r echo=FALSE, warning=FALSE, message=FALSE}

setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP_Code_Reviewed/PerisYague2023/Raw_Results")
emorating<-read.csv('EmotionalRatings_preproc.csv')
emorating<-as.data.frame(emorating)

emorating$subject <- as.factor(emorating$subject)
emorating$age <- as.numeric(emorating$age)
emorating$gender <- as.factor(emorating$gender)
emorating$word <- as.factor(emorating$word)
emorating$salience <- as.factor(emorating$salience)
emorating$list <- as.factor(emorating$list)
emorating$rating <- as.numeric(emorating$rating)

emorating <- droplevels(emorating)

demographics <- c(27,61,27,27,27,26,28,28,35,21,26,24,38,32,22)
demographics.mean <- mean(demographics) # mean age 44 
demographics.sd <- sd(demographics) #sd 24.04 

data  <- emorating %>%
    group_by(subject, salience) %>%
    get_summary_stats(rating, type = 'mean_ci')

data %>%
  pairwise_t_test(mean~salience, paired=TRUE, p.adjust.method='fdr')

data %>%
  cohens_d(mean~salience, paired=TRUE)

plt <- ggboxplot(data, x='salience', y = 'mean')

words  <- emorating %>%
    group_by(word) %>%
    get_summary_stats(rating, type = 'mean_sd')

path_out<-'/Raw_Results/'
#write.csv(words,paste(path_out,'emotional_ratings_mean.csv',sep = ''))


```

# Table S7- Semantic ratings 

``` {r echo=FALSE, warning=FALSE, message=FALSE}

setwd("/Users/albaperis/OneDrive - UAM/PhDUPM_/Von Restroff WP3/Paper_github/Odd_SOA_CRP_Code_Reviewed/PerisYague2023/Raw_Results")
semrating<-read.csv('SemanticRelatedness_preproc.csv')
semrating<-as.data.frame(semrating)

semrating$subject <- as.factor(semrating$subject)
semrating$age <- as.numeric(semrating$age)
semrating$gender <- as.factor(semrating$gender)
semrating$word <- as.factor(semrating$word)
semrating$list <- as.factor(semrating$list)
semrating$oddtype <- as.factor(semrating$oddtype)
semrating$rating <- as.numeric(semrating$rating)

semrating <- droplevels(semrating)

demographics <- c(24,62,60,28,27,30,27,21,26,27,32,27,31,24)
demographics.mean <- mean(demographics) # mean age 44 
demographics.sd <- sd(demographics) #sd 24.04 

data  <- semrating %>%
    group_by(subject, oddtype) %>%
    get_summary_stats(rating, type = 'mean_ci')

data <-droplevels(data)

semrating_aov<-anova_test(data=data, dv=mean, wid=subject,
                 within=c(oddtype), effect.size = 'pes'
                 )

get_anova_table(semrating_aov) 

data %>%
  pairwise_t_test(mean~oddtype, paired=TRUE, p.adjust.method='fdr')

data %>%
  cohens_d(mean~oddtype, paired=TRUE)

plt <- ggboxplot(data, x='oddtype', y = 'mean')

words  <- semrating %>%
    group_by(word) %>%
    get_summary_stats(rating, type = 'mean_sd')

path_out<-'/Raw_Results/'
#write.csv(words,paste(path_out,'semantic_ratings_mean.csv',sep = ''))

```

